---
title: "underserviced_areas"
author: "Darya Akimova"
date: "August 23, 2018"
output: 
  html_document:
    toc: true
    toc_float:
      smooth_scroll: false
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Main question to address: What areas of NYC are under-serviced by ADA accessible stations?

Plan: Somehow randomly "sample" locations in the NYC territory in (longitude, latitude) unit pairs and then calculate distance from that point to the nearest ada-certified (accessible) subway station. Normalize to nearest (accessible or not accessible) station? 

Need:
- Data for which stations are ADA-accessible
- Data for station locations (longitude, latitude)
- Some way to sample the territory of NYC (it's an odd shape)

# Setup

## Packages

```{r packages, comment=NA}
library(tidyverse)
library(ggmap)
library(sf)
library(ggthemes)
```


## Data

```{r data, comment=NA}
nyc.census.map <- st_read("./data/nyct2010_18a/nyct2010.shp")
nyc.neigh.map <- st_read("./data/nynta_18d/nynta.shp")
subway.ent.exit <- read_csv("./data/2018_update/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
subway.by.line <- read_csv("./data/2018_update/nyc_subway_stations_grouped.csv")
num.stat.by.route <- read_csv("./data/2018_update/nyc_subway_num_stat_by_line.csv")
```


```{r preview, comment=NA}
plot(nyc.census.map$geometry)
# subway entrances/exit data:
glimpse(subway.ent.exit)
# subway summary data
glimpse(subway.by.line)
glimpse(num.stat.by.route)
```


# Preliminary analysis and light cleanup

First things first, cleanup the column names of the two subway info dataframes by converting the words to lower-case and by replacing empty spaces with `_`:

```{r comment=NA}
colnames(subway.ent.exit) <- colnames(subway.ent.exit) %>% 
  str_to_lower() %>% 
  str_replace_all(" ", "_")
colnames(subway.ent.exit)
colnames(subway.by.line) <- colnames(subway.by.line) %>% 
  str_to_lower() %>% 
  str_replace_all(" ", "_")
colnames(subway.by.line)
```

Much better! 

Now for some cleanup of the subway entrance/exit dataset:

* Create a unique station name column by combining the division. line, and station_name
* Gather the route names into one column and get rid of all of the missing values


```{r}
sub.line.tidy <- subway.by.line %>% 
  separate(lines, into = c(paste("route", 1:4, sep = "_")), remove = FALSE) %>% 
  gather("route_num", "route_name", route_1:route_4) %>% 
  filter(!is.na(route_name)) %>% 
  select(-route_num) %>% 
  arrange(route_name)
glimpse(sub.line.tidy)


subway.ent.exit %>% 
  select(route1:route11) %>% 
  gather("route_num", "route_name") %>% 
  filter(!(is.na(route_name))) %>% 
  select(-route_num) %>% 
  distinct() %>% 
  anti_join(sub.line.tidy, by = "route_name")



length(unique(subway.ent.exit$station_name))
# NYC considers 472 to be the number of subway stations in the city, 424 if connected stations are considered single stations
nrow(subway.ent.exit %>% select(station_name, station_latitude:station_longitude) %>% distinct())

nrow(subway.ent.exit %>% select(division:station_name) %>% distinct())
sub.ext.cln <- subway.ent.exit %>% 
  mutate_at(vars(division:station_name), str_to_lower) %>% 
  unite("stat_name", division:station_name, sep = "_") %>% 
  mutate_at(vars(route1:route11), str_to_upper) %>% 
  distinct()
length(unique(sub.ext.cln$stat_name))
  



```



```{r}
sub.sml.test <- sub.ext.cln %>% 
  select(stat_name:station_longitude, entrance_latitude:entrance_longitude, route1:route11, entrance_type:entry, ada, free_crossover) %>%
  replace_na(
    list(
      route1 = "", route2 = "", route3 = "", route4 = "", 
      route5 = "", route6 = "", route7 = "", route8 = "", 
      route9 = "", route10 = "", route11 = ""
      )
    ) %>% 
  unite("routes", route1:route11, sep = "") %>% 
  distinct()
table(sub.sml.test$free_crossover)

sub.sml.test %>% 
  select(stat_name, free_crossover) %>% 
  distinct() %>% 
  group_by(stat_name) %>% 
  count() %>% 
  filter(n > 1)
sub.sml.test %>% 
  select(stat_name, free_crossover) %>% 
  distinct() %>% 
  filter(!free_crossover)
sub.sml.test %>% 
  select(stat_name, free_crossover, ada) %>% 
  distinct() %>% 
  group_by(ada, free_crossover) %>% 
  count()
subway.ent.exit %>% 
  filter(ada & !free_crossover)

percent_ada <- sub.sml.test %>% 
  group_by(stat_name) %>% 
  summarize(
    num_entry = n(),
    num_ada = sum(ada),
    percent_ada = round(num_ada * 100 / num_entry, 2)
  )
hist(percent_ada$percent_ada)
table(percent_ada$percent_ada)
table(sub.sml.test$entrance_type)

sub.sml.test %>% 
  group_by(station_latitude, station_longitude) %>% 
  summarize(
    num_entry = n(),
    num_ada = sum(ada),
    percent_ada = round(num_ada * 100 / num_entry, 2)
  ) %>% 
  {table(.$percent_ada)}
sub.sml.test %>% 
  group_by(station_latitude, station_longitude) %>% 
  summarize(
    num_entry = n(),
    num_free_cross = sum(free_crossover),
    percent_free_cross = round(num_free_cross * 100 / num_entry, 2)
  ) %>% 
  {table(.$percent_free_cross)}
sub.sml.test %>% 
  group_by(entrance_type, ada) %>% 
  count() %>% 
  ggplot(aes(entrance_type, n, fill = ada)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  scale_fill_tableau()
sub.sml.test %>% 
  group_by(entrance_type, ada) %>% 
  count() %>% 
  ggplot(aes(entrance_type, n, fill = ada)) +
  geom_bar(stat = "identity", position = "fill") +
  theme_minimal() +
  scale_fill_tableau()
sub.sml.test %>% 
  group_by(entrance_type, free_crossover) %>% 
  count() %>% 
  ggplot(aes(entrance_type, n, fill = free_crossover)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_minimal() +
  scale_fill_tableau()



```


```{r comment=NA}
# note that the sta
length(unique(subway.ent.exit$station_name))
nrow(subway.ent.exit %>% select(station_name, station_latitude:station_longitude) %>% distinct())
nrow(subway.ent.exit %>% select(division:station_name) %>% distinct())
sub.ext.cln %>% 
  select(stat_name, ada, ada_notes) %>% 
  filter(!(is.na(ada_notes))) %>% 
  distinct()

sub.stat.coord <- sub.ext.cln %>% 
  select(stat_name:station_longitude) %>% 
  distinct() %>% 
  group_by(stat_name) %>% 
  summarize(
    avg_stat_lat = mean(station_latitude),
    avg_stat_long = mean(station_longitude)
  )

sub.ext.sml <- sub.ext.cln %>% 
  select(stat_name, route1:route11, ada, ada_notes) %>% 
  inner_join(sub.stat.coord, by = "stat_name") %>% 
  distinct() %>% 
  gather("route_num", "route_name", route1:route11) %>% 
  filter(!is.na(route_name)) %>% 
  select(-route_num) %>% 
  distinct()
glimpse(sub.ext.sml)
sapply(sub.ext.sml, anyNA)
length(unique(sub.ext.sml$stat_name))
sub.ext.sml %>% 
  select(stat_name, route_name) %>% 
  distinct() %>% 
  nrow()

sub.ext.sml %>% 
  select(stat_name, ada, ada_notes) %>% 
  distinct() %>% 
  filter(!(is.na(ada_notes)))
sub.ext.sml %>% 
  filter(stat_name == "ind_8 avenue_world trade center")

sub.ext.sml %>%
  filter(str_detect(stat_name, "ind")) %>% 
  separate(stat_name, into = c("division", "line", "station_name"), sep = "_") %>% 
  {table(.$line)}
sub.ext.sml %>%
  filter(str_detect(stat_name, "irt")) %>% 
  separate(stat_name, into = c("division", "line", "station_name"), sep = "_") %>% 
  {table(.$line)}
sub.ext.sml %>% 
  filter(route_name == "G") %>% 
  ggplot(aes(avg_stat_long, avg_stat_lat)) + 
  geom_point() + 
  coord_quickmap()
# G train missing station stops between Smith 9th St and Church Ave, where the service 
# should add those in - follows the F train here, get the slice of stations: 
sub.ext.sml %>% 
  filter(route_name == "F") %>% 
  arrange(avg_stat_lat) %>% 
  filter(avg_stat_lat > 40.62976 & avg_stat_lat < 40.68030)
sub.ext.sml %>% 
  filter(route_name == "F") %>% 
  arrange(avg_stat_lat) %>% 
  filter(avg_stat_lat > 40.62976 & avg_stat_lat < 40.68030 & str_detect(stat_name, "4")) %>% 
  ggplot(aes(avg_stat_long, avg_stat_lat)) +
  geom_point()
sub.ext.sml %>% 
  filter(route_name == "F") %>% 
  arrange(avg_stat_lat) %>% 
  filter(avg_stat_lat > 40.62976 & avg_stat_lat < 40.68030) %>% 
  ggplot(aes(avg_stat_long, avg_stat_lat)) +
  geom_point() +
  coord_quickmap()

# G train is considered IND - grab that 4th ave station
sub.ext.sml.rt.fix <- sub.ext.sml %>% 
  bind_rows(
    sub.ext.sml %>% 
      filter(route_name == "F") %>% 
      arrange(avg_stat_lat) %>% 
      filter(avg_stat_lat > 40.63612 & avg_stat_lat < 40.67358 & str_detect(stat_name, "ind")) %>% 
      mutate(route_name = "G")
)

nyc.map.4boro <- nyc.census.map %>%  
  filter(BoroName != "Staten Island") %>% 
  mutate(BoroName = str_to_lower(BoroName))
nyc.map.4boro %>% filter(str_detect(NTAName, "park") | str_detect(NTAName, "Airport")) %>% ggplot() +geom_sf()
nyc.map.4boro %>% 
  filter(!(str_detect(NTAName, "park") | str_detect(NTAName, "Airport"))) %>% 
  ggplot(aes(Shape_Area)) +
  geom_histogram()
nyc.map.4boro.crs.geo <- st_transform(nyc.map.4boro, crs = "+init=epsg:4326")

sub.ext.sml.rt.fix %>% 
  filter(route_name == "F" | route_name == "G") %>% 
  ggplot(aes(avg_stat_long, avg_stat_lat, color = route_name)) +
  geom_point(size = 2, alpha = 0.8) +
  coord_quickmap()
nyc.map.4boro.crs.geo %>% 
  ggplot() +
  geom_sf() +
  geom_point(
    data = sub.ext.sml.rt.fix %>% 
      filter(route_name == "F" | route_name == "G"),
    aes(avg_stat_long, avg_stat_lat, color = route_name),
    size = 2, alpha = 0.8
    ) +
  theme_minimal()
nyc.map.4boro.crs.geo %>% 
  ggplot() +
  geom_sf() +
  geom_point(
    data = sub.ext.sml.rt.fix %>% 
      filter(route_name == "F"),
    aes(avg_stat_long, avg_stat_lat, color = route_name),
    size = 2, alpha = 0.8
    ) +
  theme_minimal()




sub.ext.sml.rt.fix %>% 
  select(stat_name, ada, ada_notes) %>% 
  distinct() %>% 
  filter(!(is.na(ada_notes)))


# GS and S routes are redundant - refer to the same thing, the grand central shuttle between times square and grand central

# solution - filter out "S" route, change GS to ada = FALSE
```

To check:

* ind_8 avenue_50th st / TRUE / Southbound Only - Leave as ada = TRUE (no changes)
* ind_8 avenue_world trade center / TRUE / Construction - ada = TRUE for E only 
* ind_archer av_sutphin blvd-archer av - jfk / TRUE / Check - ada = TRUE according to wikipedia page (no change)
* bmt_broadway_49th st / TRUE / Northbound Only - Leave as ada = TRUE (no change)
* bmt_broadway_times square-42nd st / TRUE / Shuttle not ADA - set S to ada = FALSE
* bmt_broadway_union square / TRUE / Lex not ADA - set 4/5/6 to ada = FALSE   
* bmt_canarsie_union square / TRUE / Lex not ADA - same as above      
* ind_concourse_kingsbridge rd / FALSE / in planning - switch to TRUE because it is complete      
* irt_lexington_23rd st / FALSE In Planning - switch to TRUE because it is complete
* irt_lexington_brooklyn bridge-city hall / TRUE / J Z not ADA - set J/Z to ada = FALSE
* irt_lexington_canal st / TRUE / Bway Nass not ADA - Only set 6 ada  = TRUE
* irt_pelham_hunts point av / FALSE / in planning - complete, set ada = TRUE     


* ind_queens boulevard_forest hills-71st av / FALSE / in planning - complete, set ada = TRUE

Note: This dataset has not been updated for the BMT Broadway Line / Yellow Line / NQRW train changes, namely it still includes the old Q route that went into Queens, and does not include the 3 new stations that were opened along 2nd Ave in Manhattan, all of which are ADA-Accessible.

```{r comment = NA}
sub.ext.ada.updt <- sub.ext.sml.rt.fix %>% 
  # change world trade center station ada
  mutate(ada = ifelse((stat_name == "ind_8 avenue_world trade center" & route_name != "E"), FALSE, ada)) %>% 
  # change times square shuttle ada, filter out extra "S" route
  filter(route_name != "S") %>% 
  mutate(ada = ifelse(route_name == "GS", FALSE, ada)) %>% 
  # change 4/5/6 at union square to ada = FALSE
  mutate(ada = ifelse(
      (
        (stat_name == "bmt_broadway_union square" | stat_name == "bmt_canarsie_union square") &
        (route_name == "4" | route_name == "5" | route_name == "6")
      ), FALSE, ada)) %>% 
  # change kingsbridge rd ada = TRUE
  mutate(ada = ifelse(stat_name == "ind_concourse_kingsbridge rd", TRUE, ada)) %>% 
  # change Lex / 23rd St stop to TRUE
  mutate(ada = ifelse(stat_name == "irt_lexington_23rd st", TRUE, ada)) %>% 
  # change J/Z at Brooklyn Bridge / City Hall to ada = FALSE
  mutate(ada = ifelse(
    (stat_name == "irt_lexington_brooklyn bridge-city hall" & (route_name == "J" | route_name == "Z")),
    FALSE, ada)) %>% 
  # change irt_lexington_canal st to ada = TRUE for 6 train only
  mutate(ada = ifelse(
    (stat_name == "irt_lexington_canal st" & route_name != "6"), FALSE, ada
  )) %>% 
  # change rt 6 hunts point av to ada = TRUE
  mutate(ada = ifelse(stat_name == "irt_pelham_hunts point av", TRUE, ada)) %>% 
  # convert the forst hills / 71st ave station to ada = TRUE for all lines
  mutate(ada = ifelse(stat_name == "ind_queens boulevard_forest hills-71st av", TRUE, ada)) %>% 
  select(-ada_notes) %>% 
  distinct()
dim(sub.ext.sml.rt.fix)
dim(sub.ext.ada.updt)
table(sub.ext.ada.updt$ada)

sub.ext.ada.updt %>% 
  filter(route_name == "GS")
sub.ext.ada.updt %>% 
  filter(route_name == "S")
percent.ada.updt <- sub.ext.ada.updt %>% 
  group_by(stat_name) %>% 
  summarize(
    num_entry = n(),
    num_ada = sum(ada),
    percent_ada = round(num_ada * 100 / num_entry, 2)
  )
hist(percent.ada.updt$percent_ada)
table(percent.ada.updt$percent_ada)
percent.ada.updt %>% 
  group_by(num_entry) %>% 
  count() %>% 
  ggplot(aes(num_entry, n)) +
  geom_bar(stat = "identity")
percent.ada.updt %>%
  group_by(num_ada) %>% 
  count() %>% 
  ggplot(aes(num_ada, n)) +
  geom_bar(stat = "identity")
```


This process brought up an additional problem - lines are assigned to routes that they do not belong in, and there are extra stops (probably because some stations are connected and technically the entrances/exits apply to those routes as well).

```{r comment= NA}
stat.by.route.count <- sub.ext.ada.updt %>% 
  group_by(route_name) %>% 
  count()  
stat.by.route.count %>%  
  ggplot(aes(route_name, n)) +
  geom_bar(stat = "identity")
stat.route.join <- stat.by.route.count %>% 
  full_join(num.stat.by.route, by = "route_name") %>% 
  filter(route_name != "W")
stat.route.join %>% 
  ggplot(aes(num_stations_norm, n)) +
  geom_abline(intercept = 0, slope = 1, size = 1, alpha = 0.8) +
  geom_point(size = 2, alpha = 0.8) +
  xlab("stat num by route (wiki)") +
  ylab("stat num by route (count)")
stat.route.join %>% 
  filter(num_stations_norm == n)
```


# Cleaning each route

```{r data_clean_by_route, comment=NA}
num.stat.by.route %>% 
  filter(route_name == "GS")
sub.ext.ada.updt %>% 
  filter(route_name == "GS") %>% 
  arrange(stat_name)
sub.stat.num.updt <- sub.ext.ada.updt %>% 
  filter(!(route_name == "GS" & (str_detect(stat_name, "flushing") | str_detect(stat_name, "lexington"))))
num.stat.by.route %>% 
  filter(route_name == "FS")
stat.by.route.count %>% 
  filter(route_name == "FS")
sub.stat.num.updt %>% 
  filter(route_name == "FS") %>% 
  arrange(avg_stat_lat)
# remove the ind and irt stations - those are the extras
sub.stat.num.updt <- sub.stat.num.updt %>% 
  filter(!(route_name == "FS" & (str_detect(stat_name, "irt") | str_detect(stat_name, "ind"))))
sub.stat.num.updt %>% 
  filter(route_name == "FS")
sub.stat.num.updt %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>% 
  filter(num_stations_norm == min(num_stations_norm))
# G route fix
sub.line.tidy %>% 
  filter(route_name == "G")

sub.stat.num.updt %>% 
  filter(route_name == "G") %>% 
  select(avg_stat_lat, avg_stat_long) %>% 
  distinct() %>% 
  nrow()

sub.stat.num.updt.G <- sub.stat.num.updt %>% 
  filter(route_name != "G") %>% 
  bind_rows(
    sub.stat.num.updt %>% 
      filter(route_name == "G" & str_detect(stat_name, "ind") & stat_name != "ind_queens boulevard_23rd st-ely av")
    )
sub.stat.num.updt.G %>% 
  filter(route_name == "G") %>% 
  nrow()
# Z route fix
sub.line.tidy %>% 
  filter(route_name == "Z")
sub.stat.num.updt.Z <- sub.stat.num.updt.G %>% 
  filter(route_name != "Z") %>% 
  bind_rows(
    sub.stat.num.updt.G %>% 
      filter(route_name == "Z" & str_detect(stat_name, "bmt") & stat_name != "bmt_broadway_canal st (ul)") %>% 
      # Z train was not included on the list of trains at broadway junction stop in Queens (was A/C/J/L)
      bind_rows(
        sub.stat.num.updt.G %>% 
          filter(str_detect(stat_name, "broadway junction")) %>% 
          mutate(route_name = "Z") %>% 
          distinct()
        ) %>% 
      # Z train was also not listed on Alabama Ave stop in Queens (was J only)
      bind_rows(
        sub.stat.num.updt.G %>% 
          filter(str_detect(stat_name, "alabama")) %>% 
          mutate(route_name = "Z")
      ) %>% 
      # add back in the jamaica center and jfk airport stops that were filtered out
      bind_rows(
        sub.stat.num.updt.G %>% 
          filter(route_name == "Z" & avg_stat_long > -73.82829)
        )
    )
sub.stat.num.updt.Z %>% 
  filter(route_name == "Z") %>% 
  nrow()

## Next fix targets ##
sub.stat.num.updt.Z %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>% 
  filter(num_stations_norm == min(num_stations_norm))

# 7 Train, goal: 22
sub.line.tidy %>% 
  filter(route_name == "7")
sub.stat.num.updt.7 <-  sub.stat.num.updt.Z %>%
  filter(route_name != "7") %>% 
  bind_rows(
    sub.stat.num.updt.Z %>% 
      filter(route_name == "7" & str_detect(stat_name, "irt")) %>% 
      # eliminate station copies
      filter(
        # times square extra
        stat_name != "irt_42nd st shuttle_times square" &
        # grand central extra
          stat_name != "irt_42nd st shuttle_grand central" &
          stat_name != "irt_lexington_grand central-42nd st"
          ) %>% 
      mutate(ada = ifelse(stat_name == "irt_flushing_45 rd-court house sq", TRUE, ada))
  )
sub.stat.num.updt.7 %>% 
  filter(route_name == "7") %>% 
  nrow()
# E train, goal: 22
sub.line.tidy %>% 
  filter(route_name == "E")
sub.stat.num.updt.E <- sub.stat.num.updt.7 %>% 
  filter(route_name != "E") %>% 
  bind_rows(
    sub.stat.num.updt.7 %>% 
      filter(route_name == "E" & str_detect(stat_name, "ind")) %>% 
      filter(stat_name != "ind_8 avenue_chambers st") %>% 
      mutate(ada = ifelse(stat_name == "ind_archer av_jamaica-van wyck", TRUE, ada)) %>% 
      bind_rows(
        # add in Briarwood station (a late night station)
        sub.stat.num.updt.7 %>% 
          filter(str_detect(stat_name, "briarwood")) %>% 
          mutate(route_name = "E")
        )
    )

## Next fix targets ##
sub.stat.num.updt.E %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>% 
  filter(num_stations_norm == min(num_stations_norm))

# L train, goal: 24
sub.line.tidy %>% 
  filter(route_name == "L")
sub.stat.num.updt.L <-  sub.stat.num.updt.E %>% 
  filter(route_name != "L") %>% 
  bind_rows(
    sub.stat.num.updt.E %>% 
      filter(route_name == "L" & str_detect(stat_name, "bmt")) %>% 
      filter(stat_name != "bmt_broadway_union square") %>% 
      mutate(ada = ifelse(stat_name == "bmt_canarsie_wilson av", TRUE, ada)) %>% 
      # add back in broadway junction
      bind_rows(
        sub.stat.num.updt.E %>% 
          filter(str_detect(stat_name, "broadway junction") & route_name == "L")
        )
    )
sub.stat.num.updt.L %>% 
  filter(route_name == "L") %>% 
  nrow()
## Next fix targets ##
sub.stat.num.updt.L %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>% 
  filter(num_stations_norm == min(num_stations_norm))

# B line: target - 37 - will include rush hour stations 
sub.line.tidy %>% 
  filter(route_name == "B")
sub.stat.num.updt.B <- sub.stat.num.updt.L %>% 
  filter(route_name != "B") %>% 
  bind_rows(
    sub.stat.num.updt.L %>% 
      filter(route_name == "B" & (str_detect(stat_name, "bmt") | str_detect(stat_name, "ind"))) %>% 
      # convert non-ada stations to those that are ada = TRUE now
      mutate(
        ada = ifelse((
          stat_name == "bmt_brighton_kings highway" | 
            stat_name == "ind_6 avenue_broadway-lafayette st" | 
            stat_name == "ind_8 avenue_125th st"
          ), TRUE, ada)
        ) %>% 
      # stations to exclude atlantic ave /barclays duplicates 
      # and stops between barclays and brighton where B does not stop
      filter(!(avg_stat_lat > 40.60867 & avg_stat_lat < 40.63508)) %>% 
      filter(!(
        stat_name %in% c(
          "bmt_broadway_34th st", "bmt_brighton_parkside av", 
          "bmt_4 avenue_pacific st", "bmt_brighton_atlantic av", 
          "bmt_brighton_av u", "bmt_brighton_neck rd", 
          "bmt_brighton_beverly rd", "bmt_brighton_cortelyou rd"
          )
        ))
    )
dim(sub.stat.num.updt.B)

## Next fix targets ##
sub.stat.num.updt.B %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B") %>% 
  filter(num_stations_norm == min(num_stations_norm))

# 4 train, target is 28 - exclude late night service
sub.stat.num.updt.4 <- sub.stat.num.updt.B %>% 
  filter(route_name != "4") %>% 
  bind_rows(
    sub.stat.num.updt.B %>% 
      filter(route_name == "4" & str_detect(stat_name, "irt")) %>% 
      filter(!(
        stat_name %in% c(
          "irt_flushing_grand central-42nd st", "irt_42nd st shuttle_grand central", 
          "irt_clark_fulton st", "irt_clark_borough hall"
          )
        )) %>% 
      mutate(ada = ifelse(stat_name == "irt_lexington_fulton st", TRUE, ada))
    )

# J train, goal: 30
sub.stat.num.updt.J <- sub.stat.num.updt.4 %>% 
  filter(route_name != "J") %>% 
  bind_rows(
    sub.stat.num.updt.4 %>% 
      filter(route_name == "J" & str_detect(stat_name, "bmt")) %>%
      filter(stat_name != "bmt_broadway_canal st (ul)") %>% 
      bind_rows(
        sub.stat.num.updt.4 %>% 
        filter(route_name == "J" & avg_stat_long > -73.82829)
      ) %>% 
      bind_rows(
        sub.stat.num.updt.4 %>% 
              filter(str_detect(stat_name, "broadway junction") & route_name == "J")
      ) %>% 
      mutate(ada = ifelse(stat_name == "bmt_nassau_fulton st", TRUE, ada))
    )

sub.stat.num.updt.J %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")

# route A, goal: 44 stations
sub.stat.num.updt.A <- sub.stat.num.updt.J %>% 
  filter(route_name != "A") %>% 
  bind_rows(
    sub.stat.num.updt.J %>% 
      filter(route_name == "A" & str_detect(stat_name, "ind")) %>%
      filter(!(
        stat_name %in% c(
          "ind_8 avenue_world trade center", "ind_8 avenue_broadway-nassau", 
          "ind_fulton_franklin av", "ind_fulton_kingston-throop",
          "ind_fulton_ralph av", "ind_fulton_rockaway av",
          "ind_fulton_liberty av", "ind_fulton_van siclen av", 
          "ind_fulton_shepherd av"
          )
        )) %>% 
      bind_rows(
        sub.stat.num.updt.J %>% 
          filter(str_detect(stat_name, "fulton st") & route_name == "4") %>% 
          mutate(route_name = "A")
      ) %>% 
      mutate(ada = ifelse(stat_name %in% c("ind_8 avenue_125th st", "ind_rockaway_far rockaway-mott av", "ind_rockaway_aqueduct racetrack", "ind_fulton_jay st - borough hall", "ind_fulton_utica av", "ind_liberty_lefferts blvd"), TRUE, ada)) %>% 
      bind_rows(
        sub.stat.num.updt.J %>% 
          filter(route_name == "H" & !(str_detect(stat_name, "broad channel"))) %>% 
          mutate(route_name = "A")
      )
    )
nrow(sub.stat.num.updt.A)

sub.stat.num.updt.A %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")
# C next, goal: 40 stations
sub.stat.num.updt.C <- sub.stat.num.updt.A %>% 
  filter(route_name != "C") %>% 
  bind_rows(
    sub.stat.num.updt.A %>% 
      filter(route_name == "C" & str_detect(stat_name, "ind")) %>% 
      filter(!(stat_name %in% c("ind_8 avenue_broadway-nassau", "ind_8 avenue_world trade center"))) %>% 
      bind_rows(
            sub.stat.num.updt.A %>% 
              filter(str_detect(stat_name, "fulton st") & route_name == "A") %>% 
              mutate(route_name = "C")
        ) %>%
      mutate(ada = ifelse(stat_name %in% c("ind_8 avenue_125th st", "ind_fulton_jay st - borough hall", "ind_fulton_utica av"), TRUE, ada))
)
sub.stat.num.updt.C %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")
# route 5, goal: 45 
sub.stat.num.updt.5 <- sub.stat.num.updt.C %>% 
  filter(route_name != "5") %>% 
  bind_rows(
    sub.stat.num.updt.C %>% 
      filter(route_name == "5" & str_detect(stat_name, "irt")) %>% 
      filter(!(
        stat_name %in% c(
          "irt_clark_borough hall", "irt_clark_fulton st", 
          "irt_flushing_grand central-42nd st", "irt_42nd st shuttle_grand central", 
          "irt_white plains road_wakefield-241st st"
          )
        )) %>%
      mutate(ada = ifelse(
        stat_name %in% c("irt_lexington_fulton st", "irt_white plains road_east 180th st", "irt_white plains road_gun hill rd"), TRUE, ada
        ))
    )
sub.stat.num.updt.5 %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")

# Next: route 6, goal = 38
sub.stat.num.updt.6 <- sub.stat.num.updt.5 %>% 
  filter(route_name != "6") %>% 
  bind_rows(
    sub.stat.num.updt.5 %>% 
      filter(route_name == "6" & str_detect(stat_name, "irt")) %>% 
      filter(!(stat_name %in% c("irt_flushing_grand central-42nd st", "irt_42nd st shuttle_grand central"))) %>% 
      mutate(ada = ifelse(stat_name %in% c("irt_lexington_bleecker st"), TRUE, ada))
    )
sub.stat.num.updt.6 %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")
# Next: route D, goal: 36
sub.stat.num.updt.D <- sub.stat.num.updt.6 %>% 
  filter(route_name != "D") %>% 
  bind_rows(
    sub.stat.num.updt.6 %>% 
      filter(route_name == "D" & (str_detect(stat_name, "ind") | str_detect(stat_name, "bmt"))) %>% 
      filter(!(
        stat_name %in% c(
          "bmt_broadway_34th st", "bmt_4 avenue_pacific st", 
          "bmt_brighton_atlantic av", "bmt_sea beach_new utrecht av", 
          "bmt_brighton_stillwell av")
        )) %>% 
      bind_rows(
        sub.stat.num.updt.6 %>% 
          filter(str_detect(stat_name, "4 avenue_36")) %>% 
          mutate(route_name = "D") %>% 
          distinct()
        ) %>% 
      mutate(ada = ifelse(
        stat_name %in% c("ind_8 avenue_125th st", "ind_6 avenue_broadway-lafayette st", "bmt_west end_bay parkway"), TRUE, ada)
        )
    )

sub.stat.num.updt.D %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")

# Next: route F, goal: 45
sub.stat.num.updt.F <- sub.stat.num.updt.D %>% 
  filter(route_name != "F") %>% 
  bind_rows(
    sub.stat.num.updt.D %>% 
      filter(route_name == "F" & (str_detect(stat_name, "ind") | str_detect(stat_name, "bmt"))) %>% 
      filter(!(
        stat_name %in% c(
          "bmt_broadway_34th st", "bmt_canarsie_6th av", 
          "bmt_nassau_essex st", "bmt_broadway_lawrence st", 
          "bmt_4 avenue_9th st", "bmt_brighton_stillwell av", 
          "bmt_brighton_west 8th st"
          )
        )) %>% 
      mutate(ada = ifelse(stat_name %in% c("ind_6 avenue_broadway-lafayette st", "ind_fulton_jay st - borough hall"), TRUE, ada))
    )
nrow(sub.stat.num.updt.F %>% filter(route_name == "F"))
sub.stat.num.updt.F %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")
# Next: route M, goal: 36
sub.stat.num.updt.M <- sub.stat.num.updt.F %>% 
  filter(route_name != "M") %>% 
  bind_rows(
    sub.stat.num.updt.F %>% 
      filter(route_name == "M" & (str_detect(stat_name, "ind") | str_detect(stat_name, "bmt"))) %>% 
      filter(!(stat_name %in% c("bmt_broadway_34th st", "bmt_canarsie_6th av", "bmt_nassau_essex st"))) %>% 
      mutate(ada = ifelse(stat_name %in% c("ind_6 avenue_broadway-lafayette st"), TRUE, ada))
    )
sub.stat.num.updt.M %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")
# Next: route 1, goal: 38
sub.stat.num.updt.1 <- sub.stat.num.updt.M %>% 
  filter(route_name != "1") %>% 
  bind_rows(
    sub.stat.num.updt.M %>% 
      filter(route_name == "1" & str_detect(stat_name, "irt")) %>%
      filter(!(stat_name %in% c("irt_42nd st shuttle_times square", "irt_clark_park place"))) %>% 
      # add in the re-opened WTC Cortlandt station (coord from wikipedia)
      bind_rows(
        tibble(
          stat_name = "irt_broadway-7th ave_wtc cortlandt", ada = TRUE, 
          avg_stat_lat = 40.7115, avg_stat_long = -74.012, route_name = "1")
        ) %>% 
      mutate(
        ada = ifelse(
          stat_name %in% c(
            "irt_broadway-7th ave_dyckman st", "irt_broadway-7th ave_168th st"
            ), 
          TRUE, ada))
    )
sub.stat.num.updt.1 %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")
# Next: route 2, goal: 49 or 52?
sub.stat.num.updt.2 <- sub.stat.num.updt.1 %>% 
  filter(route_name != "2") %>% 
  bind_rows(
    sub.stat.num.updt.1 %>% 
      filter(route_name == "2" & str_detect(stat_name, "irt")) %>% 
      filter(!(
        stat_name %in% c(
          "irt_42nd st shuttle_times square", "irt_lexington_fulton st",
          "irt_lexington_borough hall"
          )
        )) %>% 
      mutate(
        ada = ifelse(stat_name %in% c(
          "irt_white plains road_gun hill rd", "irt_white plains road_east 180th st",
          "irt_clark_fulton st"
          ), TRUE, ada)
        )
    )
sub.stat.num.updt.2 %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")

#Next: route 3, goal: 34
sub.stat.num.updt.3 <- sub.stat.num.updt.2 %>% 
  filter(route_name != "3") %>% 
  bind_rows(
    sub.stat.num.updt.2 %>% 
      filter(route_name == "3" & str_detect(stat_name, "irt")) %>% 
      filter(!(stat_name %in% c(
        "irt_42nd st shuttle_times square", "irt_lexington_fulton st", "irt_lexington_borough hall"
        ))) %>%
      mutate(ada = ifelse(stat_name %in% c("irt_clark_fulton st"), TRUE, ada))
    )
sub.stat.num.updt.3 %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")
# What I've been dreading, R goal = 45
sub.stat.num.updt.R <- sub.stat.num.updt.3 %>% 
  filter(route_name != "R") %>% 
  bind_rows(
    sub.stat.num.updt.3 %>% 
      filter(route_name == "R" & (str_detect(stat_name, "bmt") | str_detect(stat_name, "ind"))) %>% 
      filter(!(stat_name %in% c(
        "ind_6 avenue_smith-9th st", "bmt_4 avenue_pacific st", 
        "bmt_brighton_atlantic av", "ind_fulton_jay st - borough hall", 
        "bmt_nassau_canal st", "bmt_canarsie_union square", "ind_6 avenue_34th st", "ind_8 avenue_42nd st"
        ))) %>% 
      mutate(ada = ifelse(stat_name %in% c("bmt_broadway_lawrence st", "bmt_broadway_cortlandt st"), TRUE, ada))
    )
sub.stat.num.updt.R %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")
# Next: route N, num stations - 32
sub.stat.num.updt.N <- sub.stat.num.updt.R %>% 
  filter(route_name != "N") %>% 
  bind_rows(
    sub.stat.num.updt.R %>% 
      filter(route_name == "N" & str_detect(stat_name, "bmt")) %>% 
      filter(!(stat_name %in% c(
        "bmt_brighton_stillwell av", "bmt_west end_62nd st", 
        "bmt_brighton_atlantic av", "bmt_4 avenue_pacific st", 
        "bmt_nassau_canal st", "bmt_canarsie_union square"
        ))) %>% 
      bind_rows(
        sub.stat.num.updt.R %>% filter(route_name == "N" & str_detect(stat_name, "queensboro"))
      )
    )
sub.stat.num.updt.N %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")

num.stat.by.route %>% 
  filter(route_name == "W")

sub.stat.num.updt.W <- sub.stat.num.updt.N %>% 
  bind_rows(
    sub.stat.num.updt.N %>% 
      filter(route_name == "N" & avg_stat_lat > 40.68367) %>% 
      bind_rows(
        sub.stat.num.updt.N %>% 
          filter(route_name == "R") %>% 
          filter(avg_stat_lat > 40.69410 & avg_stat_lat < 40.71952)
        ) %>% 
      mutate(route_name = "W")
    )
sub.stat.num.updt.W %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")
# last route: Q, with 29 stations

sub.stat.num.updt.final <- sub.stat.num.updt.W %>% 
  filter(route_name != "Q") %>% 
  bind_rows(
    sub.stat.num.updt.W %>% 
      filter(route_name == "Q" & !str_detect(stat_name, "astoria") & str_detect(stat_name, "bmt")) %>%
      bind_rows(
        tibble(
          stat_name = c("ind_2 avenue_72nd st", "ind_2 avenue_86th st", "ind_2 avenue_96th st"), 
          ada = rep(TRUE, 3), 
          avg_stat_lat = c(40.768889, 40.777861, 40.7841), 
          avg_stat_long = c(-73.958333, -73.95175, -73.9472), 
          route_name = rep("Q", 3)
          )
        ) %>% 
      bind_rows(
        sub.stat.num.updt.W %>% 
          filter(stat_name == "ind_63rd street_lexington av") %>% 
          mutate(route_name = "Q")
      ) %>% 
      filter(!(stat_name %in% c(
        "bmt_broadway_5th av", "bmt_broadway_lexington av", 
        "bmt_broadway_49th st", "bmt_canarsie_union square", 
        "bmt_nassau_canal st", "bmt_4 avenue_pacific st", 
        "bmt_brighton_atlantic av", "bmt_coney island_stillwell av", 
        "bmt_coney island_west 8th st"
        ))) %>%
      mutate(ada = ifelse(stat_name %in% c("bmt_brighton_av h", "bmt_brighton_kings highway"), TRUE, ada))
    )
sub.stat.num.updt.final %>% 
  group_by(route_name) %>% 
  count() %>% 
  ungroup() %>% 
  inner_join(num.stat.by.route, by = "route_name") %>% 
  filter(n != num_stations_norm) %>%
  filter(route_name != "B")
```


# Analysis and Visualization

```{r comment=NA}
nyc.map.4boro <- nyc.census.map %>%  
  filter(BoroName != "Staten Island") %>% 
  mutate(BoroName = str_to_lower(BoroName))
tract.cent <- st_centroid(nyc.map.4boro)

sub.cln.sf <- st_as_sf(sub.stat.num.updt.final, coords = c("avg_stat_long", "avg_stat_lat"), crs = 4326)
sub.sf.nyc.crs <- st_transform(sub.cln.sf, crs = st_crs(nyc.map.4boro))

plot(st_geometry(nyc.map.4boro), col = "grey", border = "white")
plot(sub.sf.nyc.crs, pch = 16, cex = 0.25, col = "#D55E00", add = TRUE)

sub.boro <- st_join(sub.sf.nyc.crs, nyc.map.4boro)
```


```{r}
theme_set(theme_minimal())
sub.boro.nogeo <- st_set_geometry(sub.boro, NULL) %>% 
  as_tibble() %>% 
  select(stat_name:route_name, BoroName:BoroCT2010, NTAName, Shape_Area)
stat.by.route <- sub.boro.nogeo %>% 
  group_by(route_name) %>% 
  summarize(
    total_stat = n(),
    num_ada = sum(ada),
    per_ada = num_ada * 100 / total_stat
  ) %>% 
  mutate(
    rt_mod = ifelse(
      route_name == "GS" | route_name == "FS" | route_name == "H",
      "S",
      route_name
      )
    ) %>% 
  left_join(sub.line.tidy, by = c("rt_mod" = "route_name")) %>% 
  arrange(primary_trunk_line, route_name) 
stat.by.route$rt_order <- factor(stat.by.route$route_name, levels = stat.by.route$route_name)

subway.by.line <- subway.by.line %>% 
  arrange(primary_trunk_line)

stat.by.route %>% 
  ggplot(aes(rt_order, total_stat, fill = primary_trunk_line)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = subway.by.line$hexadecimal, name = "Line") +
  ylab("Total Number of Stations") +
  xlab("Subway Route")
stat.by.route %>% 
  ggplot(aes(rt_order, num_ada, fill = primary_trunk_line)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = subway.by.line$hexadecimal, name = "Line") +
  ylab("Number of ADA Stations") +
  xlab("Subway Route")
stat.by.route %>% 
  ggplot(aes(rt_order, per_ada, fill = primary_trunk_line)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = subway.by.line$hexadecimal, name = "Line") +
  ylab("Percent of Stations that are ADA") +
  xlab("Subway Route")

stat.by.route <- stat.by.route %>% 
  arrange(total_stat) %>% 
  mutate(total_stat_order = factor(route_name, levels = route_name))
 
stat.by.route %>% 
  ggplot(aes(total_stat_order, total_stat)) +
  geom_segment(aes(x = total_stat_order, xend = total_stat_order, y = num_ada, yend = total_stat), size = 1, alpha= 0.5, color = "#56B4E9") +
  geom_point(size = 3, color = "#0072B2") +
  geom_point(aes(total_stat_order, num_ada), size = 3, color = "#56B4E9") +
  xlab("Route Name") +
  coord_flip()

stat.by.route %>% 
  ggplot(aes(total_stat, num_ada, color = primary_trunk_line)) +
  geom_abline(slope = 1, intercept = 0, alpha = 0.5, size = 1.5) +
  geom_point(size = 3) +
  scale_color_manual(values = subway.by.line$hexadecimal, name = "Line") +
  xlab("Total Number of Stations") +
  ylab("Number of ADA Stations") +
  theme_minimal()
stat.by.route %>% 
  ggplot(aes(total_stat, per_ada, color = primary_trunk_line)) +
  geom_point(size = 3) +
  scale_color_manual(values = subway.by.line$hexadecimal, name = "Line") +
  xlab("Total Number of Stations") +
  ylab("Percent of Stations that are ADA") +
  theme_minimal()
stat.by.route %>% 
  ggplot(aes(total_stat, per_ada, color = primary_trunk_line, label = route_name)) +
  geom_point(size = 3) +
  scale_color_manual(values = subway.by.line$hexadecimal, name = "Line") +
  geom_text(hjust = 0, vjust = 0, check_overlap = TRUE, nudge_y = 1.5) +
  xlab("Total Number of Stations") +
  ylab("Percent of Stations that are ADA") +
  theme_minimal()
```


```{r boro_plots, comment=NA}
sub.boro.nogeo

boro.stat.count <- sub.boro.nogeo %>%
  group_by(BoroName) %>% 
  summarize(
    tot_stat = n(),
    ada_stat = sum(ada),
    ada_percent = ada_stat * 100 / tot_stat
    )
  
boro.stat.count %>% 
  select(-ada_percent) %>% 
  gather("stat_type", "num_stat", tot_stat:ada_stat) %>% 
  ggplot(aes(BoroName, num_stat, fill = stat_type)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_tableau()
boro.stat.count %>% 
  ggplot(aes(BoroName, ada_percent, fill = BoroName)) + 
  geom_bar(stat= "identity") +
  scale_fill_tableau()
neigh.stat.count <- sub.boro.nogeo %>% 
  group_by(BoroName, NTAName) %>% 
  summarize(
    tot_stat = n(),
    ada_stat = sum(ada),
    ada_percent = ada_stat * 100 / tot_stat
    )
library(ggridges)
neigh.stat.count %>% 
  select(-ada_percent) %>% 
  gather("stat_type", "num_stat", tot_stat:ada_stat) %>% 
  ggplot(aes(y = BoroName, x = num_stat, fill = stat_type)) +
  geom_density_ridges()
neigh.stat.count %>% 
  arrange(desc(tot_stat))
neigh.stat.count %>% 
  ggplot(aes(y = BoroName, x = ada_percent, fill = BoroName)) +
  geom_density_ridges() +
  scale_fill_tableau()
neigh.stat.count %>% 
  ggplot(aes(x = BoroName, y = ada_percent, color = BoroName)) +
  geom_jitter(size = 2, alpha = 0.8, width = 0.25) +
  scale_color_tableau()

```


Imitate report: https://comptroller.nyc.gov/reports/service-denied-accessibility-and-the-new-york-city-subway-system/

```{r}
nyc.neigh.4boro <- nyc.neigh.map %>% 
  filter(BoroName != "Staten Island")
plot(st_geometry(nyc.neigh.4boro))
nyc.neigh.stat.count <- st_join(sub.sf.nyc.crs, nyc.neigh.4boro) %>% 
 st_set_geometry(NULL) %>% 
  group_by(BoroName, NTAName) %>% 
  summarize(
    tot_stat = n(),
    ada_stat = sum(ada),
    percent_ada = ada_stat * 100 / tot_stat
  ) %>% 
  ungroup()
glimpse(nyc.neigh.stat.count)
nyc.neigh.stat.count %>% 
  ggplot(aes(BoroName, tot_stat, color = BoroName)) +
  geom_jitter(size = 2, alpha = 0.8, width = 0.25) +
  scale_color_tableau()
data_summary <- function(x) {
   m <- mean(x)
   ymin <- ifelse(m - sd(x) > 0, m - sd(x), 0)
   ymax <- m + sd(x)
   return(c(y=m,ymin=ymin,ymax=ymax))
}
nyc.neigh.stat.count %>% 
  ggplot(aes(BoroName, tot_stat, color = BoroName, fill = BoroName)) +
  geom_violin() +
  stat_summary(fun.data = data_summary, color = "black", alpha = 0.8) +
  scale_color_tableau() +
  scale_fill_tableau() +
  xlab("Borough") +
  ylab("Subway Stations per Neighborhood")
nyc.neigh.stat.count %>% 
  ggplot(aes(BoroName, ada_stat, color = BoroName, fill = BoroName)) +
  geom_violin() +
  stat_summary(fun.data = data_summary, color = "black", alpha = 0.8) +
  scale_color_tableau() +
  scale_fill_tableau() +
  xlab("Borough") +
  ylab("Subway Stations per Neighborhood")

nyc.neigh.4boro.count <- nyc.neigh.4boro %>% 
  full_join(
    nyc.neigh.stat.count %>% 
      select(-BoroName),
    by = "NTAName"
    )
nyc.neigh.4boro.count %>% 
  ggplot() +
  geom_sf(aes(fill = tot_stat)) 
nyc.neigh.count.prep <- nyc.neigh.4boro.count %>% 
  mutate(
    color_group = ifelse(
      str_detect(NTAName, "park"), "Park", ifelse(
        str_detect(NTAName, "Airport"), "Airport", ifelse(
          !(str_detect(NTAName, "park") | str_detect(NTAName, "Airport")) & is.na(tot_stat), "No stations", ifelse(
            !(str_detect(NTAName, "park") | str_detect(NTAName, "Airport")) & tot_stat > 0 & ada_stat < 1, "No accessible stations", ifelse(
              !(str_detect(NTAName, "park") | str_detect(NTAName, "Airport")) & tot_stat > 0 & ada_stat > 0, "At least one accessible station", "Don't know"
              )
            )
          )
        )
      )
    )
table(nyc.neigh.count.prep$color_group)
nyc.neigh.count.prep %>% 
  ggplot() +
  geom_sf(aes(fill = color_group), color = "#BFBFBF") +
  scale_fill_manual(values = c("grey30", "#0072B2", "#E69F00", "#999999", "#bae4b3")) +
  geom_sf(data = sub.sf.nyc.crs, color = "#BFBFBF", fill = "black", size = 1, pch = 21)
nyc.neigh.count.prep %>% 
  ggplot() +
  geom_sf(aes(fill = tot_stat), color = "#BFBFBF") +
  scale_fill_viridis_c(option = "D")
nyc.neigh.count.prep %>% 
  ggplot() +
  geom_sf(aes(fill = ada_stat), color = "#BFBFBF") +
  scale_fill_viridis_c(option = "D")
nyc.neigh.count.prep$ada_per_breaks <- cut(nyc.neigh.count.prep$percent_ada, breaks = seq(0, 100, 20))
nyc.neigh.count.prep %>% 
  ggplot() +
  geom_sf(aes(fill = ada_per_breaks), color = "#BFBFBF") +
  scale_fill_manual(values = c("#c7e9b4", "#7fcdbb", "#41b6c4", "#2c7fb8", "#253494"), na.value = "gray50")
summary(nyc.neigh.count.prep$percent_ada)
```



```{r census_tract_buffer, comment=NA}
tract.cent <- st_centroid(nyc.map.4boro)
nyc.map.4boro %>% 
  ggplot() +
  geom_sf(color = "#BFBFBF") +
  geom_sf(data = tract.cent, size = 1, color = "#4E79A7")
# half-mile buffer
tract.cent.halfmi.buff <- st_buffer(tract.cent, dist = 2640)
nyc.map.4boro %>% 
  ggplot() +
  geom_sf(color = "#BFBFBF") +
  geom_sf(data = tract.cent.halfmi.buff, alpha = 0.2, color = "#F28E2B") +
  geom_sf(data = tract.cent, size = 1, color = "#4E79A7")
#example
nyc.map.4boro %>% 
  ggplot() +
  geom_sf(color = "#BFBFBF") +
  geom_sf(data = tract.cent.halfmi.buff %>% filter(BoroCT2010 == 1009800 | BoroCT2010 == 1018400 ), alpha = 0.8, color = "#F28E2B", fill = "#F28E2B") +
  geom_sf(data = sub.sf.nyc.crs, color = "black", size = 1)



  



sub.sf.nyc.crs %>% 
  inner_join(
    sub.boro %>% 
      st_set_geometry(NULL) %>% 
      as_tibble() %>% 
      select(stat_name, BoroName) %>% 
      distinct(),
    by = "stat_name"
  ) %>% 
  st_join(
    tract.cent.halfmi.buff
    )



sub.buffer.join <- st_join(sub.sf.nyc.crs, tract.cent.halfmi.buff) %>% 
  st_set_geometry(NULL)

nyc.map.4boro.test <- nyc.map.4boro %>% 
  full_join(
    sub.buffer.join %>% 
      group_by(BoroCT2010) %>% 
      count() %>% 
      ungroup(),
    by = "BoroCT2010"
    )
length(unique(test3$BoroCT2010))
length(unique(nyc.map.4boro$BoroCT2010))

nyc.map.4boro.test %>% 
  ggplot() +
  geom_sf(aes(fill = n)) +
  scale_fill_viridis_c(option = "A") +
  geom_sf(data = sub.sf.nyc.crs, color = "white", size = 1)
```


# Session Info

```{r session_info, comment=NA}
sessionInfo()
```

